---
title: "Final Project"
author: "Luke Friedman"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

incsv = "C:/Users/Luke Friedman/Documents/Polypterus/bluered/bluereddistrays.csv"
 ndata <- read.csv(incsv, header = TRUE)
print(head(ndata$x))
print(sum(is.na(as.numeric(ndata$x))))
library(tidyverse)
library(ggplot2)
library(stringr)
```

Scan Intensity to Bone Density Regressions
```{r}
#A3HAS <- read.csv("C:/Users/Luke Friedman/Documents/Polypterus/A3/A3HAS.csv")
regressionHAS <- function(file_path) {
  d1 <- read.csv(file_path, header = TRUE)
  colnames(d1) <- c('group', 'values')
  d1$values <- as.numeric(d1$values)
  n_group <- nrow(d1) %/% 5
  
  means <- sapply(1:n_group, function(i) {mean(d1$values[((i-1)*5+1):(i*5)])})
  print(means)
  
  ex_den <- c(-20.379, 101.8616, 212.4351, 417.3955, 789.2673)
  matrix <- matrix(c(means, ex_den), nrow=2, byrow=TRUE)

rownames(matrix) <- c("Intensity Measured", "HA Vals")
print(matrix)

plot(means, ex_den, ylab="Density Standards", xlab="Measured Intensity")
mod <- lm(ex_den ~ means)
abline(mod)
print(summary(mod))
}
lm1 <- regressionHAS("C:/Users/Luke Friedman/Documents/Polypterus/B5/HASB5.csv")
lm1
cf <- coef(lm1)
Intercept1 <- cf[1]
Slope1 <- cf[2]

intenseregression <- function(incsv, outcsv, m, b) {
  ndata <- read.csv(incsv, header = TRUE)
  ndata$y <- m * ndata$x + b
  write.csv(ndata, outcsv, row.names = FALSE)
  print(paste("New data:", outcsv))
}
intenseregression(
  incsv = "C:/Users/Luke Friedman/Documents/Polypterus/B5/B5proxrays.csv", 
  outcsv = "B5proxraysnew.csv", 
  m = Slope1, 
  b = Intercept1)

A2.1densities <- read.csv('B5proxraysnew.csv')
View(A2.1densities)
write.csv(A2.1densities,"~/Polypterus/B5/B5proxraysnew.csv", row.names = FALSE)

```
Data Conglomeration to Single CSV File
```{r}
path_to_data <- "C:/Users/Luke Friedman/Documents/Polypterus"

specimen_ids <- list( Aquatic = c("A2", "A3", "B4", "B5", "C2"), Terrestrial = c("bluered", "gror", "orgwht", "redblue", "redred"))


# Bone groups
bone_groups <- list(
  Radials = list(filename_suffix = "Radials.csv"),
    
  Proximal_Fin_Rays = list(filename_suffix = "proxrays.csv"),
  Distal_Fin_Rays = list(filename_suffix = "distrays.csv"),
  Prop_Metap = list(filename_suffix = "PrMt.csv")
)


process_csv <- function(file_path, bone_group, specimen_id, treatment_group = list()) {
  data <- read.csv(file_path)
  
  if ("Density" %in% colnames(data)) {
    bone_density <- data[["Density"]]
  } else {
    bone_density <- data[[3]]
  }
  
  if (is.null(bone_density) || length(bone_density) == 0) {
    warning(paste("No valid bone density data found in file:", file_path))
    return(NULL)
  }
  
  processed_data <- data.frame(
    Specimen.ID = specimen_id,
    Treatment.Group = treatment_group,
    Bone.Group = bone_group,
    BoneDensity = bone_density
  )
  
  return(processed_data)
}


all_data <- data.frame()




for (treatment in names(specimen_ids)) {
  for (specimen in specimen_ids[[treatment]]) {
    for (bone_group in names(bone_groups)) {
      file_path <- file.path(path_to_data, paste0(specimen, "_", bone_groups[[bone_group]]$filename_suffix))
      
      if (!file.exists(file_path)) {
        warning(paste("File not found:", file_path)) }
      
        bone_data <- process_csv(
          file_path = file_path,
          bone_group = bone_group,
          specimen_id = specimen,
          treatment_group = treatment
        )
      
          all_data <- rbind(all_data, bone_data)
        }
      } 
      }



output_file_path <- file.path(path_to_data, "consolidated_bone_density_data1.csv")
write.csv(all_data, file = output_file_path, row.names = FALSE)
cat("Final data saved to:", output_file_path, "\n")
```
Initial Visualization

```{r}
data <- read.csv("consolidated_bone_density_data.csv")
library(tidyverse)



hist1 <- data[data$Treatment.Group %in% c("Aquatic", "Terrestrial"),]


ggplot(data=hist1, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + labs(title="Bone Density by Treatment Group", x="Bone Density (mg/ccm)")

ggplot(data=hist1, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)")

ggplot(data = hist1, aes(x = BoneDensity, fill = Treatment.Group)) + geom_histogram(position = "dodge", binwidth = 100)

ggplot(data=hist1, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)") + facet_wrap(~Bone.Group)

library(stringr)

dorven <- hist1 %>%
  mutate(Position_Group = case_when(
    str_detect(Position, regex("Dorsal|Proximal1|Proximal2|Proximal3|Proximal4|Proximal5|Distal1|Distal2|Distal3|Distal4|Distal5|Mid1|Mid2|Mid3|Mid4|Mid5|propterygium")) ~ "Dorsal",
    str_detect(Position, regex("Ventral|Proximal6|Proximal7|Proximal8|Proximal9|Proximal10|Distal6|Distal7|Distal8|Distal9|Distal10|Mid6|Mid7|Mid8|Mid9|Mid10|metapterygium")) ~ "Ventral",
    TRUE ~ "Other"
  ))

ggplot(data=dorven, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)") + facet_wrap(~Position_Group)









```
Identification of Outlying Specimens
```{r}
ggplot(data=hist1, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group", x="Bone Density (mg/ccm)") +facet_wrap(~Specimen.ID) + coord_flip()

static <- data %>% group_by(Treatment.Group, Specimen.ID) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity))

ggplot(static, aes(x=Specimen.ID, y=meandensity, color=Treatment.Group)) + geom_point(size=2)+ geom_errorbar(aes(ymin=meandensity-SDdensity, ymax=meandensity+SDdensity)) + labs(title = "Mean Bone Density and Standard Deviation by Individual", x = "Specimen", y = "Bone Density (mg/ccm)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Filtered Visualizations
```{r}
filtdata <- data %>% filter(!Specimen.ID %in% c("bluered", "C2", "B4"))

hist2 <- filtdata[filtdata$Treatment.Group %in% c("Aquatic", "Terrestrial"),]

ggplot(data=hist2, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + labs(title="Bone Density by Treatment Group", x="Bone Density (mg/ccm)") #+ xlim(0, 2000)

ggplot(data=hist2, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)")

ggplot(data=hist2, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)") + facet_wrap(~Bone.Group)

ggplot(data=hist2, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)") +facet_wrap(~Specimen.ID) + coord_flip()


dorven2 <- hist2 %>%
  mutate(Position_Group = case_when(
    str_detect(Position, regex("Dorsal|Proximal1|Proximal2|Proximal3|Proximal4|Proximal5|Distal1|Distal2|Distal3|Distal4|Distal5|Mid1|Mid2|Mid3|Mid4|Mid5|propterygium")) ~ "Dorsal",
    str_detect(Position, regex("Ventral|Proximal6|Proximal7|Proximal8|Proximal9|Proximal10|Distal6|Distal7|Distal8|Distal9|Distal10|Mid6|Mid7|Mid8|Mid9|Mid10|metapterygium")) ~ "Ventral",
    TRUE ~ "Other"
  ))

ggplot(data=dorven2, aes(x=BoneDensity, fill=Treatment.Group)) + geom_density(alpha=0.5) + scale_x_log10() + labs(title="Bone Density by Treatment Group (Log-Scale)", x="Bone Density (mg/ccm)") + facet_wrap(~Position_Group) + xlim(0,2000)

static2 <- filtdata %>% group_by(Treatment.Group, Specimen.ID) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity))

ggplot(static2, aes(x=Specimen.ID, y=meandensity, color=Treatment.Group)) + geom_point(size=2)+ geom_errorbar(aes(ymin=meandensity-SDdensity, ymax=meandensity+SDdensity)) + labs(title = "Mean Bone Density and Standard Deviation by Individual", x = "Specimen", y = "Bone Density (mg/ccm)")

static3 <- data %>% group_by(Treatment.Group) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity))

ggplot(static3, aes(x=Treatment.Group, y=meandensity)) + geom_point(size=2)+ geom_errorbar(aes(ymin=meandensity-SDdensity, ymax=meandensity+SDdensity)) + labs(title = "Mean Bone Density and Standard Deviation by Individual", x = "Specimen", y = "Bone Density (mg/ccm)")

static4 <- filtdata %>% group_by(Treatment.Group) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity))

ggplot(static4, aes(x=Treatment.Group, y=meandensity)) + geom_point(size=2)+ geom_errorbar(aes(ymin=meandensity-SDdensity, ymax=meandensity+SDdensity)) + labs(title = "Mean Bone Density and Standard Deviation by Individual", x = "Specimen", y = "Bone Density (mg/ccm)")

```
Simple T-Test for Overall Bone Densities
```{r}
ttestoverall1 <- t.test(BoneDensity ~ Treatment.Group, data=hist1, var.equal=TRUE)
print(ttestoverall1)
ttestoverall2 <- t.test(BoneDensity ~ Treatment.Group, data=hist2, var.equal=TRUE)
print(ttestoverall2)
```
Resampling Trials to Observe Robustness of Data
```{r}
get_p_value_t_test <- function() {
  resamp <- hist1 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}
set.seed(123)
pvalues <- replicate(n = 1000, expr=get_p_value_t_test())
sig <- sum(pvalues < 0.05)
hist(pvalues, main="Unfiltered for Errant Specimens")
print(sig)
pvaluesadj <- p.adjust(pvalues, method = "bonferroni")
sig2 <- sum(pvaluesadj < 0.05)
print(sig2)

get_p_value_t_test2 <- function() {
  resamp <- hist2 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}
set.seed(123)
pvalues <- replicate(n = 1000, expr=get_p_value_t_test2())
sig3 <- sum(pvalues < 0.05)
hist(pvalues, main="Filtered for Errant Specimens")
print(sig3)
pvaluesadj <- p.adjust(pvalues, method = "bonferroni")
sig4 <- sum(pvaluesadj < 0.05)
print(sig4)
```
T-Tests by Bone Group (Filtered and Unfiltered Data)
```{r}
boneavg <- data %>% group_by(Bone.Group, Treatment.Group) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity), count=n())
print(boneavg)

ttest1 <- data %>% group_by(Bone.Group) %>% summarise(t_value = t.test(BoneDensity[Treatment.Group == "Aquatic"], BoneDensity[Treatment.Group == "Terrestrial"])$statistic, pvalues= t.test(BoneDensity[Treatment.Group == "Aquatic"], BoneDensity[Treatment.Group == "Terrestrial"])$p.value)
print(ttest1)

boneavg2 <- filtdata %>% group_by(Bone.Group, Treatment.Group) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity), count=n())

print(boneavg2)

ttest2 <- filtdata %>% group_by(Bone.Group) %>% summarise(t_value = t.test(BoneDensity[Treatment.Group == "Aquatic"], BoneDensity[Treatment.Group == "Terrestrial"])$statistic, pvalues= t.test(BoneDensity[Treatment.Group == "Aquatic"], BoneDensity[Treatment.Group == "Terrestrial"])$p.value)
print(ttest2)
```

Resampling Trials By Bone Groups
```{r}
get_p_value_t_test3 <- function(bonegroup) {
  resamp <- hist1 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}
pvals_by_group <- list()
set.seed(123)
for (bone_group in unique(hist1$Bone.Group)) {
    
  bonegroup <- hist1 %>% filter(Bone.Group == bone_group)
  pvalues <- replicate(n = 1000, expr=get_p_value_t_test3(bonegroup))
  pvals_by_group[[bone_group]] <- pvalues
}

hist(pvals_by_group[[1]], main=paste("p-values for", names(pvals_by_group)[1]))
hist(pvals_by_group[[2]], main=paste("p-values for", names(pvals_by_group)[2]))
hist(pvals_by_group[[3]], main=paste("p-values for", names(pvals_by_group)[3]))
hist(pvals_by_group[[4]], main=paste("p-values for", names(pvals_by_group)[4]))


sig2_by_bone_group <- sapply(pvals_by_group, function(pvals) sum(pvals < 0.05))

print(sig2_by_bone_group)


get_p_value_t_test4 <- function(bonegroup) {
  resamp <- hist2 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}
pvals_by_group <- list()
set.seed(123)
for (bone_group in unique(hist2$Bone.Group)) {
    
  bonegroup <- hist2 %>% filter(Bone.Group == bone_group)
  pvalues <- replicate(n = 1000, expr=get_p_value_t_test4(bonegroup))
  pvals_by_group[[bone_group]] <- pvalues
}

hist(pvals_by_group[[1]], main=paste("p-values for", names(pvals_by_group)[1]))
hist(pvals_by_group[[2]], main=paste("p-values for", names(pvals_by_group)[2]))
hist(pvals_by_group[[3]], main=paste("p-values for", names(pvals_by_group)[3]))
hist(pvals_by_group[[4]], main=paste("p-values for", names(pvals_by_group)[4]))


sig2_by_bone_group <- sapply(pvals_by_group, function(pvals) sum(pvals < 0.05))

print(sig2_by_bone_group)
```
Bonferroni Adjustment Counts by Bone Group
```{r}
p_value_t_test3 <- function(bonegroup) {
  resamp <- hist1 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}


pvals_by_group <- list()
set.seed(123)
for (bone_group in unique(hist1$Bone.Group)) {
  bonegroup <- hist1 %>% filter(Bone.Group == bone_group)
  pvalues <- replicate(n = 1000, expr=p_value_t_test3(bonegroup))
  pvals_by_group[[bone_group]] <- pvalues
}


all_pvals <- unlist(pvals_by_group)


adjusted_pvals <- p.adjust(all_pvals, method = "bonferroni")


adjusted_pvals_by_group <- split(adjusted_pvals, rep(1:length(pvals_by_group), times = sapply(pvals_by_group, length)))


sig2_by_bone_group <- sapply(adjusted_pvals_by_group, function(pvals) sum(pvals < 0.05))

print(sig2_by_bone_group)


p_value_t_test4 <- function(bonegroup) {
  resamp <- hist2 %>% sample_frac(1, replace = TRUE)
  tresults <- t.test(BoneDensity ~ Treatment.Group, data = resamp, var.equal = TRUE)
  return(tresults$p.value)
}


pvals_by_group <- list()
set.seed(123)
for (bone_group in unique(hist2$Bone.Group)) {
  bonegroup <- hist2 %>% filter(Bone.Group == bone_group)
  pvalues <- replicate(n = 1000, expr=p_value_t_test4(bonegroup))
  pvals_by_group[[bone_group]] <- pvalues
}

all_pvals <- unlist(pvals_by_group)

adjusted_pvals <- p.adjust(all_pvals, method = "bonferroni")

adjusted_pvals_by_group <- split(adjusted_pvals, rep(1:length(pvals_by_group), times = sapply(pvals_by_group, length)))

sig2_by_bone_group <- sapply(adjusted_pvals_by_group, function(pvals) sum(pvals < 0.05))

print(sig2_by_bone_group)

```

Linear Regression (Dorsal-Ventral Analysis) of Radials
```{r}
rad1 <- data %>% filter(Bone.Group == "Radials") %>% group_by(Specimen.ID) %>% mutate(Radnum = rep(1:15, each=5, length.out=n())) %>% group_by(Specimen.ID, Radnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(rad1)

ggplot(rad1, aes(x=factor(Radnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Radials (Log-Scale)", y="Bone Density (mg/ccm)") + scale_y_log10() 

rad2 <- filtdata %>% filter(Bone.Group == "Radials") %>% group_by(Specimen.ID) %>% mutate(Radnum = rep(1:15, each=5, length.out=n())) %>% group_by(Specimen.ID, Radnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(rad2)

ggplot(rad2, aes(x=factor(Radnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Radials", y="Bone Density (mg/ccm)")

ggplot(rad2, aes(x=Radnum, y=meanbonedense)) + geom_point() + geom_smooth(method="lm", color='blue') + labs(title="Bone Density of Radials", y="Bone Density (mg/ccm)")

lmrad <- lm(meanbonedense ~ Radnum, data = rad2)
summary(lmrad)

radialavg <- rad2 %>% group_by(Radnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(radialavg, aes(x=Radnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)")

lmrad2 <- lm(avgbonedense ~ Radnum, data = radialavg)
summary(lmrad2)

radavgfiltered <- rad2 %>% group_by(Radnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE)) %>% filter(Radnum !=15)

ggplot(radavgfiltered, aes(x=Radnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(title="Bone Density of Radials", y="Bone Density (mg/ccm)")
lmrad3 <- lm(avgbonedense ~ Radnum, data = radavgfiltered)
summary(lmrad3)

rad3 <- data %>% filter(Bone.Group == "Radials") %>% group_by(Specimen.ID) %>% mutate(Radnum = rep(1:15, each=5, length.out=n())) %>% group_by(Specimen.ID, Radnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

radavgunfiltered <- rad3 %>% group_by(Radnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE)) %>% filter(Radnum !=15)

ggplot(radavgunfiltered, aes(x=Radnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(title="unfiltered", y="Bone Density (mg/ccm)")
lmrad4 <- lm(avgbonedense ~ Radnum, data = radavgunfiltered)
summary(lmrad4)
```
Linear Regression (Dorsal-Ventral Analysis) of Prox Fin Rays
```{r}
prox1 <- data %>% filter(Bone.Group == "Proximal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Proxnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Proxnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(prox1)

ggplot(prox1, aes(x=factor(Proxnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Proximal Rays (Log-Scale)", y="Bone Density (mg/ccm)") + scale_y_log10() 

prox2 <- filtdata %>% filter(Bone.Group == "Proximal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Proxnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Proxnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(prox2)

ggplot(prox2, aes(x=factor(Proxnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Proximal Rays", y="Bone Density (mg/ccm)")

ggplot(prox2, aes(x=Proxnum, y=meanbonedense)) + geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)")

lmprox <- lm(meanbonedense ~ Proxnum, data = prox2)
summary(lmprox)

proxavg <- prox2 %>% group_by(Proxnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(proxavg, aes(x=Proxnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)")

lmprox2 <- lm(avgbonedense ~ Proxnum, data = proxavg)
summary(lmprox2)

prox3 <- data %>% filter(Bone.Group == "Proximal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Proxnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Proxnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

proxavg <- prox3 %>% group_by(Proxnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(proxavg, aes(x=Proxnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)")

lmprox3 <- lm(avgbonedense ~ Proxnum, data = proxavg)
summary(lmprox3)
```
Linear Regression (Dorsal-Ventral Analysis) of Distal Fin Rays
```{r}
dist1 <- data %>% filter(Bone.Group == "Distal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Distnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Distnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(dist1)

ggplot(dist1, aes(x=factor(Distnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Distal Rays (Log-Scale)", y="Bone Density (mg/ccm)") + scale_y_log10() 

dist2 <- filtdata %>% filter(Bone.Group == "Distal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Distnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Distnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(dist2)

ggplot(dist2, aes(x=factor(Distnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Distal Rays", y="Bone Density (mg/ccm)")

ggplot(dist2, aes(x=Distnum, y=meanbonedense)) + geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)")

lmdist <- lm(meanbonedense ~ Distnum, data = dist2)
summary(lmdist)

distavg <- dist2 %>% group_by(Distnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(distavg, aes(x=Distnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs(y="Bone Density (mg/ccm)") 

lmdist2 <- lm(avgbonedense ~ Distnum, data = distavg)
summary(lmdist2)

dist3 <- data %>% filter(Bone.Group == "Distal_Fin_Rays") %>% group_by(Specimen.ID) %>% mutate(Distnum = rep(1:10, each=5, length.out=n())) %>% group_by(Specimen.ID, Distnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

distavg <- dist3 %>% group_by(Distnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(distavg, aes(x=Distnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue') + labs( y="Bone Density (mg/ccm)")

lmdist3 <- lm(avgbonedense ~ Distnum, data = distavg)
summary(lmdist3)

```

Linear Regression (Dorsal-Ventral Analysis) of Propterygium and Metapterygium
```{r}
PrMt1 <- data %>% filter(Bone.Group == "Prop_Metap") %>% group_by(Specimen.ID) %>% mutate(Prnum = rep(1:2, each=5, length.out=n())) %>% group_by(Specimen.ID, Prnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(PrMt1)

ggplot(PrMt1, aes(x=factor(Prnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Propterygium and Metapterygium (Log-Scale)", y="Bone Density (mg/ccm)") + scale_y_log10() 

Pr2 <- filtdata %>% filter(Bone.Group == "Prop_Metap") %>% group_by(Specimen.ID) %>% mutate(Prnum = rep(1:2, each=5, length.out=n())) %>% group_by(Specimen.ID, Prnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

print(Pr2)

ggplot(Pr2, aes(x=factor(Prnum), y=meanbonedense)) + geom_boxplot() + labs(title="Bone Density of Propterygium and Metapterygium", y="Bone Density (mg/ccm)")

ggplot(Pr2, aes(x=Prnum, y=meanbonedense)) + geom_point() + geom_smooth(method="lm", color='blue')+ labs( y="Bone Density (mg/ccm)")

lmpr <- lm(meanbonedense ~ Prnum, data = Pr2)
summary(lmpr)

pravg <- Pr2 %>% group_by(Prnum) %>% summarise(avgbonedense=mean(meanbonedense, na.rm=TRUE))
ggplot(pravg, aes(x=Prnum, y=avgbonedense))+geom_point() + geom_smooth(method="lm", color='blue')+ labs(y="Bone Density (mg/ccm)")

lmpr2 <- lm(meanbonedense ~ Prnum, data = Pr2)
summary(Pr2)


t_test_result <- t.test(meanbonedense ~ Prnum, data = Pr2, var.equal = TRUE)
print(t_test_result)


t_test_result <- t.test(meanbonedense ~ Prnum, data = PrMt1, var.equal = TRUE)
print(t_test_result)
```
Dorsal-Ventral Axis Comparisons by Treatment (and Bone Group)
```{r}
library(tidyverse)
library(ggplot2)

rad2 <- filtdata %>%
  filter(Bone.Group == "Radials") %>%
  group_by(Specimen.ID, Treatment.Group) %>% 
  mutate(Radnum = rep(1:15, each=5, length.out=n())) %>%
  group_by(Treatment.Group, Radnum) %>%
  summarise(meanbonedense = mean(BoneDensity))

ggplot(rad2, aes(x=Radnum, y=meanbonedense, color=Treatment.Group)) + geom_point() + geom_smooth(method="lm", se=TRUE) + labs(y="Mean Bone Density (mg/ccm)")

#rad2 <- filtdata %>% filter(Bone.Group == "Radials") %>% group_by(Specimen.ID) %>% mutate(Radnum = rep(1:15, each=5, length.out=n())) %>% group_by(Specimen.ID, Radnum) %>% summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))

prox2 <- filtdata %>%
  filter(Bone.Group == "Proximal_Fin_Rays") %>%
  group_by(Specimen.ID, Treatment.Group) %>%
  mutate(Proxnum = rep(1:10, each=5, length.out=n())) %>%
  group_by(Treatment.Group, Proxnum) %>%
  summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))
ggplot(prox2, aes(x = Proxnum, y = meanbonedense, color = Treatment.Group)) + geom_point() + geom_smooth(method="lm", se=TRUE) + labs(y="Mean Bone Density (mg/ccm)")

dist2 <- filtdata %>%
  filter(Bone.Group == "Distal_Fin_Rays") %>%
  group_by(Specimen.ID, Treatment.Group) %>%
  mutate(Distnum = rep(1:10, each=5, length.out=n())) %>%
  group_by(Treatment.Group, Distnum) %>%
  summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))
ggplot(dist2, aes(x = Distnum, y = meanbonedense, color = Treatment.Group)) + geom_point() + geom_smooth(method="lm", se=TRUE) + labs(y="Mean Bone Density (mg/ccm)")

Pr2 <- filtdata %>%
  filter(Bone.Group == "Prop_Metap") %>%
  group_by(Specimen.ID, Treatment.Group) %>%
  mutate(Prnum = rep(1:2, each=5, length.out=n())) %>%
  group_by(Treatment.Group, Prnum) %>%
  summarise(meanbonedense = mean(BoneDensity, na.rm = TRUE))
ggplot(Pr2, aes(x = Prnum, y = meanbonedense, color = Treatment.Group)) + geom_point() + geom_smooth(method="lm", se=TRUE) + labs(y="Mean Bone Density (mg/ccm)")

lm_radials <- lm(meanbonedense ~ Radnum * Treatment.Group, data = rad2)
summary(lm_radials)
lm_prox <- lm(meanbonedense ~ Proxnum * Treatment.Group, data = prox2)
summary(lm_prox)

lm_dist <- lm(meanbonedense ~ Distnum * Treatment.Group, data = dist2)
summary(lm_dist)

lm_pr <- lm(meanbonedense ~ Prnum * Treatment.Group, data = Pr2)
summary(lm_pr)

```

ANOVA analysis for Density by Treatment Group
```{r}
anova_results <- filtdata %>%
  group_by(Bone.Group) %>%
  summarise(
    anova_table = list(broom::tidy(aov(BoneDensity ~ Treatment.Group, data = cur_data()))),
    .groups = "drop"
  ) %>%
  unnest(anova_table)  


print(anova_results)

anova_results <- data %>%
  group_by(Bone.Group) %>%
  summarise(
    anova_table = list(broom::tidy(aov(BoneDensity ~ Treatment.Group, data = cur_data()))),
    .groups = "drop"
  ) %>%
  unnest(anova_table)  


print(anova_results)




```

Proximal-Distal Trends by Bone Group Overall
#```{r}
library(ggplot2)
with(boneavg2, table(Bone.Group, Treatment.Group))
boneavg2 <- filtdata %>% group_by(Bone.Group, Treatment.Group) %>% summarise(meandensity=mean(BoneDensity), SDdensity=sd(BoneDensity), count=n())
print(boneavg2)
boneavg2 <- droplevels(boneavg2)
boneavg2 <- boneavg2 %>%
  group_by(Bone.Group, Treatment.Group) %>%
  filter(n() > 1) %>%
  ungroup()
boneavg2$Bone.Group <- as.numeric(boneavg2$Bone.Group, levels=c("Prop_Metap", "Radials", "Proximal_Fin_Rays", "Distal_Fin_Rays"))
#boneavg2$Bone.Group <- factor(boneavg2$Bone.Group, levels = c("Prop_Metap", "Radials", "Proximal_Fin_Rays", "Distal_Fin_Rays"))
#boneavg2$Bone.Group <- as.numeric(boneavg2$Bone.Group)

ggplot(boneavg2, aes(x=Bone.Group, y=meandensity, color=Treatment.Group))+ geom_point(position= position_dodge(width=0.5), size=2)+  geom_smooth(method="lm", se=TRUE) +theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="Bone Density by Bone Group Filtered", y="Mean Bone Density (mg/ccm)")
lm_pr <- lm(meandensity ~ Bone.Group * Treatment.Group, data = boneavg2)
summary(lm_pr)

#```
#```{r}
print(table(boneavg2$Bone.Group))
print(table(boneavg2$Treatment.Group))

#```
```{r}
# Define boneavg2
boneavg2 <- filtdata %>%
  group_by(Bone.Group, Treatment.Group) %>%
  summarise(meandensity = mean(BoneDensity), 
            SDdensity = sd(BoneDensity), 
            count = n(), .groups = "drop")

# Convert Bone.Group to factor and then numeric
boneavg2$Bone.Group <- factor(boneavg2$Bone.Group, levels = c("Prop_Metap", "Radials", "Proximal_Fin_Rays", "Distal_Fin_Rays"))
boneavg2$Bone.Group <- as.numeric(boneavg2$Bone.Group)

# Drop unused levels
boneavg2 <- droplevels(boneavg2)

# ggplot
ggplot(boneavg2, aes(x = Bone.Group, y = meandensity, color = Treatment.Group)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bone Density by Bone Group Filtered", y = "Mean Bone Density (mg/ccm)")

# Linear model
lm_pr <- lm(meandensity ~ Bone.Group * Treatment.Group, data = boneavg2)
summary(lm_pr)
```


#```{r}
# Convert Bone.Group to a factor and map levels to numbers
boneavg2$Bone.Group <- factor(boneavg2$Bone.Group, levels = c("Prop_Metap", "Radials", "Proximal_Fin_Rays", "Distal_Fin_Rays"))
boneavg2$Bone.Group <- as.numeric(boneavg2$Bone.Group)

# Create the ggplot
ggplot(boneavg2, aes(x = Bone.Group, y = meandensity, color = Treatment.Group)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bone Density by Bone Group Filtered", y = "Mean Bone Density (mg/ccm)")

# Fit the linear model
lm_pr <- lm(meandensity ~ Bone.Group * Treatment.Group, data = boneavg2)
summary(lm_pr)

#```


Proximal-Distal Axis Within Distal Fin Rays
```{r}
library(tidyverse)
dist2 <- filtdata %>%
  filter(Bone.Group == "Distal_Fin_Rays") %>%
  mutate(
    ProximalDistal = rep(c("Proximal", "Mid-Proximal", "Mid", "Mid-Distal", "Distal"), length.out = n())
  )
dist_summary <- dist2 %>%
  group_by(ProximalDistal, Treatment.Group) %>%
  summarise(
    mean_density = mean(BoneDensity, na.rm = TRUE),
    se_density = sd(BoneDensity, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
ggplot(dist_summary, aes(x = ProximalDistal, y = mean_density, color = Treatment.Group)) +
  geom_point(size = 3, position = position_dodge(width = 0.2)) +
  geom_line(aes(group = Treatment.Group), position = position_dodge(width = 0.2)) +
  geom_errorbar(aes(ymin = mean_density - se_density, ymax = mean_density + se_density),
                width = 0.2, position = position_dodge(width = 0.2)) +
  labs(
    title = "Proximal-Distal Trends in Bone Density for Distal Fin Rays",
    x = "Proximal-Distal Level",
    y = "Mean Bone Density"
  )

anova_model <- aov(
  BoneDensity ~ ProximalDistal *  Treatment.Group,
  data = dist2
)
summary(anova_model)
```